---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: LINK."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
```


# Introduction

The English language has evolved a great deal over centuries and studying its evolution can give insight into the people who use it. Linguists are concerned with studying languages as a science; they observe and listen to speakers, conduct experiments, and perform analyses of languages to investigate the properties and characteristics of particular languages. Linguists may study the frequency of letter usage for a basic understanding of the particular writing system adopted by a language, whether it be syllabic, ideographic, or alphabetic. Additionally, the frequency of letters may vary from language to language and such information is of interest to cryptographers, keyboard designers, game creators.


```{r}
#| label: fig-agegroupavg
#| fig-cap: "The average of all shelter visitors in each age group over 6 years from 2018 to 2023"
#| echo: false
#| warning: false
#| message: false
```


# Data {#sec-data}

Some of our data is of penguins (@fig-bills), from @palmerpenguins.




Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)



Talk way more about it. 



# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.  

\begin{align} 
y_i|\mu_i &\sim \mbox{Poisson}(\mu_i) \\
\mbox{log}(\mu_i) &= \beta_0 + \beta_1 x Number of Words_i\\
\beta_0 &\sim \mbox{Normal}(0, 2.5) \\
\beta_1 &\sim \mbox{Normal}(0, 2.5) \\
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.


### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.


# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/frankenstein_e_counts.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

```{r}
#| echo: false
#| eval: true
#| label: fig-modelresultsfig
#| fig-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

library(marginaleffects)

plot_predictions(first_model, condition = "word_count") +
  labs(x = "Number of words",
       y = "Average number of e/Es in the first 10 lines") +
  theme_classic()
```





# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this. 

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}


# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows... 

```{r}

```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```



\newpage


# References


